Reproducible Research: Peer Assessment 1
==========================================================

## Loading and preprocessing the data

```{r, echo=TRUE}
activity <- read.csv("activity.csv", colClasses = c("numeric", "character", "numeric"))
head(activity)
```


Change the variable date (a factor variable) to a 'Date' variable
```{r, echo=TRUE}
activity$date <- as.Date(activity$date, "%Y-%m-%d")
```

## What is mean total number of steps taken per day?

Create a variable that contain total steps and plot a histogram
```{r, echo=TRUE}
TotalSteps <- aggregate(steps ~ date, data = activity, sum, na.rm = TRUE)
hist(TotalSteps$steps, main = "Total steps by day", xlab = "day", col="blue")
```

Show mean and median of of total steps taker per day
```{r, echo=TRUE}
mean(TotalSteps$steps)
median(TotalSteps$steps)
```

## What is the average daily activity pattern?

Create a variable that takes average steps per day and plot a time series graph
```{r, echo=TRUE}
AverageSteps <- tapply(activity$steps, activity$interval, mean, na.rm = TRUE)

plot(row.names(AverageSteps), AverageSteps, type = "l", xlab = "5-min interval", 
     ylab = "Average across all Days", main = "Average number of steps taken",
     col = "blue")
```

Find the maximum steps taken into one day
```{r, echo=TRUE}
MaximumSteps <- which.max(AverageSteps)
names(MaximumSteps)
```

## Imputing missing values

Fill the missing values with the mean of steps taken that day

```{r, echo=TRUE}
MissingValues <- sum(is.na(activity))
MissingValues

FillMissingValues <- aggregate(steps ~ interval, data = activity, FUN = mean)
fillNA <- numeric()
for (i in 1:nrow(activity)) {
  obs <- activity[i, ]
  if (is.na(obs$steps)) {
    steps <- subset(FillMissingValues, interval == obs$interval)$steps
  } else {
    steps <- obs$steps
  }
  fillNA <- c(fillNA, steps)
}

DatanoMissings <- activity
DatanoMissings$steps <- fillNA
```

Plot a histogram of the total number of steps taken each day and report the mean and median total number of steps taken per day

```{r, echo=TRUE}
TotalSteps2 <- aggregate(steps ~ date, data = DatanoMissings, sum, na.rm = TRUE)
hist(TotalSteps2$steps, main = "Total steps by day", xlab = "day", col = "blue")

mean(TotalSteps2$steps)
median(TotalSteps2$steps)
``` 

## Are there differences in activity patterns between weekdays and weekends?

Add new variables to the dataset 'activity', indicating the day of that measure and if it was a weekday or a weekend

```{r, echo=TRUE}
activity$Day <- weekdays(activity$date)
activity$DayType <- c("weekday")

for (i in 1:nrow(activity)){
  if (activity$Day[i] == "sábado" || activity$Day[i] == "domingo"){
    activity$DayType[i] <- "weekend"
  }
}

activity$DayType <- as.factor(activity$DayType)
```

Plot the average of steps taken in weekdays and weekends

```{r, echo=TRUE}
IntervalSteps <- aggregate(steps ~ interval+DayType, activity, mean)

library(ggplot2)

qplot(interval, steps, data=IntervalSteps, geom=c("line"), xlab="Interval", 
      ylab="Number of steps", main="") + facet_wrap(~ DayType, ncol=1)
```
